Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-23 papers/thesis-kirstie/thesis.tex
edits

\begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
    \caption{(Left) The WCA potential is a minimum at r=2R. 
             (Right) The WCA potential increases rapidly for $r<2R$.}
    \label{fig:TwoSpheres}
    \end{figure}

%MOVE to METHODS start >>>
%........................
%MOVE to METHODS  end  >>>

\section{The Theory of Freezing and Maxwell's Double Tangent Construction}

@@

%move this paragraph to methods?
Although the entire crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell_firsttime} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. The parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, includes portions of eight atoms for a total of one full 
atom, and has eight lattice sites convenietly reachable by moving along 
one of three different lattice vectors to step from one atom to the next. 
The three lattice vectors will be useful in creating the lattice structure 
computationally, and will also make the program run faster.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell_firsttime}
  \end{figure}
%-------------------------------

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

@@

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

Once the pressure is known for the transition point, the density of the 
liquid at the transition point can be found from a plot of the pressure 
versus the inverse of the density for the liquid. The density of the solid 
at the transition point can be found the same way, as shown in Figure 
\ref{fig:Pvsinvn}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{MaxwellDTC-Fig3.pdf}
    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are 
    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
can be generated by repeating this procedure for various temperatures. 
%The lowest density the crystal exhibts and the highest density the fluid 
%exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transistion. The gray region is where the liquid and the solid coexist, 
    and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz 
%free energy would be for both the liquid and solid phases at a particular 
%temperature and density. Whichever phase has the lower Helmholtz free 
%energy will be the phase that the substance will be in.

@@

Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}
  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
$dT=0$ gives
\begin{equation}
  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
\end{equation}
where conservation of energy was also applied. Fixing the volume and the 
number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
results in
\begin{equation}
  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
\end{equation}
which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
of particles are fixed, the change in the free energy of the system will 
always be negative, and thus the free energy tends toward a minimum. 
(reference schroder) This illustrates the tendancy toward minimization 
for a system near equilibrium. A general proof of minimization can be 
found in (reference simple fluids).

@@

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. %\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    %$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

  \begin{figure}[h!]
    \centering
    %\includegraphics[height=7cm]{diagrammic.pdf}
    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations 
    between individual pairs of atoms are shown in green. Interactions 
    between pairs of atoms with an atom shared in common are shown in red. 
    At low densities only the interactions shown in green are considered, 
    since atoms are not likely to be clustered.}
    \label{fig:diagrammic}
  \end{figure}

At low densities, it is not likely to find two or more atoms in close 
proximity, and so only interactions between individual pairs of spheres, 
such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
Each pair can then be described by a function of the distance between the 
two atoms in the pair. Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ 
for each pair we get 
\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
\color{black}
With a careful counting of the resulting number of such pairs, the excess 
free energy can then be written as 
\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
for a homogeneous density n=$\frac{N}{V}$ this becomes
\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
Thus the free excess energy per atom in the low-density limit is given by 
\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
where the quantity in parenthesis can be recognized 
as the second virial expanision coefficient typically called $B_2$. 
%where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
It is important that the expression for the excess Helmholtz free energy 
correctly reduces to this result in the low-density limit. 
%Although we will not be restricting ourselves to the low-density limit, 
%it is important that the theory we use correctly reduces to this same 
%result in the low-density limit. 
Outside of the low density limit, it is quite difficult to evaluate the 
excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}.

Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
is a function of the number density n. For a spatially varying number density, 
as would be expected for a crystal, the free energy is a functional of 
the number density profile $n(\vec{r})$  
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
where the ideal Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}.
To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
the temperature fixed until the free energy is minimized. The minimum free 
energy and its corresponding number density profile are then the equilibrium 
values for that temperature.

\section{Soft Fundamental Measure Theory (SFMT)}
Another way to form an expression for the excess Helmholtz free energy in 
terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
Scaled particle theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy.
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation is known as Widom's Insertion Formula given by
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
hard sphere, R, and the cavity is either filled or not. \color{black}
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result for different sizes of cavities in the fluid, 
Scaled Partical Theory arrives at an expression for the excess Helmholtz 
free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

@@

%MOVE to METHODS start2 >>>
%........................
%MOVE to METHODS end >>>

%MOVE to METHODS start3 >>>
%........................
%MOVE to METHODS end >>>

@@

%MOVE to METHODS start3 >>>

We use classical Density Functional Theory (cDFT) in conjunction 
with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
energy of a system, and the corresponding equilibrium number density profile. 
In particular, SFMT is used to form the expression for the excess free energy 
without having to take into account the interaction each sphere, or atom, 
has with every other sphere. The excess free energy is then added to the 
ideal free energy calculated analytically to find the resulting free energy 
for the system. 
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
For a spatially varying number density, as would be expected for a crystal, 
the free energy is a functional of the number density profile $n(\vec{r})$  
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
the temperature fixed until the free energy is minimized. The minimum free 
energy and its corresponding number density profile are then the equilibrium 
values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
the fluid are found, they can be compared to find the resulting state of 
the fluid. Outside the region of coexistence, the fluid will be completely 
in a crystalline state if the crystal free energy is lower than the liquid 
free energy, and it will be completely in a liquid state if the liquid 
free energy is lower than the crystal free energy.
%MOVE to METHODS end3 >>>

%MOVE start4 >>>>

@@

  \begin{figure}[h!]
    \centering
    \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell}
  \end{figure}

@@

 \begin{figure}[h!]
    \centering
    %\includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

A simplified flowchart of the program created to investigate the freezing 
of a WCA fluid is shown on %pages x-x. 
The program takes four inputs: reduced number density n*, reduced temperature t*, 
a range of Gaussian 
widths gw, and a range of fraction of vacancies fv.The number density 
profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
is calculated for each profile until the minimum free energy is found for 
that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
The process is then repeated to find the free energies at other number 
densities and temperatures. 
%MOVE end4 >>>>

@@

*****  end  stuff?  ******
\newline
\newline

**** Fideal - liquid and crystal  ****

The Liquid (homogeneous) ideal free energy $F_{Hideal}$ is computed analytically for a given $n^*$ and $t^*$, where $\sigma=1$ so that $n^*=n\sigma^3$=n, and $\epsilon=1$ so that $t^*=\frac{k_BT}{\epsilon}=k_BT$. 
%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} where

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{figs/TwoSpheres.pdf} 
    \caption{(Left) The WCA potential is a minimum at r=2R. 
             (Right) The WCA potential increases rapidly for $r<2R$.}
    \label{fig:TwoSpheres}
    \end{figure}

\newpage
\section{The Theory of Freezing and Maxwell's Double Tangent Construction}

@@

@@

%For each particular temperature, the pressure of the liquid and the 
%solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized 
for a system at fixed temperature, volume and number of particles as the 
entropy of the system and its surroundings is maximized. The negative of 
the partial derivative of the Helmholtz free energy with respect to volume V 
at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the 
number of particles, the partial derivative of the Helmholtz free energy 
with respect to volume becomes the partial derivative of the Helmholtz 
free energy per atom with respect to the inverse number density. 
\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

@@

The pressure at the transistion point can be obtained from the point of 
intersection that occurs when the Gibbs free energy per atom is plotted 
against the pressure for both the liquid and the solid at fixed temperature, 
as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
the slope and cooresponding y-intercept data collected from lines tangent 
along the curves on the the Helmholtz free energy per atom verses inverse 
density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

Once the pressure is known for the transition point, the density of the 
liquid at the transition point can be found from a plot of the pressure 
versus the inverse of the density for the liquid. The density of the solid 
at the transition point can be found the same way, as shown in Figure 
\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
can be generated by repeating this procedure for various temperatures. 
%The lowest density the crystal exhibts and the highest density the fluid 
%exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are 
    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
    the number densities of the liquid and the solid during a phase 
    transistion. The gray region is where the liquid and the solid coexist, 
    and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz 
%free energy would be for both the liquid and solid phases at a particular 
%temperature and density. Whichever phase has the lower Helmholtz free 
%energy will be the phase that the substance will be in.

@@

Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}
  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
$dT=0$ gives
\begin{equation}
  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
\end{equation}
where conservation of energy was also applied. Fixing the volume and the 
number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
results in
\begin{equation}
  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
\end{equation}
which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
of particles are fixed, the change in the free energy of the system will 
always be negative, and thus the free energy tends toward a 
minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
for a system near equilibrium. A general proof of minimization can be 
found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.

@@

\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. 
%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by
\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}

  \begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{figs/diagrammic.pdf}
    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations 
    between individual pairs of atoms are shown in green. Interactions 
    between pairs of atoms with an atom shared in common are shown in red. 
    At low densities only the interactions shown in green are considered, 
    since atoms are not likely to be clustered.}
    \label{fig:diagrammic}
  \end{figure}

At low densities, it is not likely to find two or more atoms in close 
proximity, and so only interactions between individual pairs of spheres, 
such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
Each pair can then be described by a function of the distance between the 
two atoms in the pair. 
%Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ for each pair we get 
%\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
With a careful counting of the resulting number of such pairs, the excess 
free energy can then be written~\cite{schroeder} as
%\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
%where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
%Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
%for a homogeneous density n=$\frac{N}{V}$ this becomes
%\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
%Thus the free excess energy per atom in the low-density limit is given by 
\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
in the thermodynamic limit of large N 
where the quantity in parenthesis can be recognized 
as the second virial expanision coefficient typically called $B_2$. 
%where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
It is important that the expression for the excess Helmholtz free energy 
correctly reduces to this result in the low-density limit. 
%Although we will not be restricting ourselves to the low-density limit, 
%it is important that the theory we use correctly reduces to this same 
%result in the low-density limit. 
Outside of the low density limit, it is quite difficult to evaluate the 
excess Helmholtz free energy given in Equation~\ref{eq:Fexcess-simplified}.

Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
is a function of the number density n. For a spatially varying number density, 
as would be expected for a crystal, the free energy becomes a functional of 
the number density profile $n(\vec{r})$.  
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
The functional for the ideal Helmholtz free energy is given by 
Equation~\ref{eq:Fideal-n(r)}, and the functional for the excess Helmholtz 
free energy depends on the particular fluid as defined by the spheres' 
interatomic potential V(r). To implement cDFT, the number density profile 
$n(\vec{r})$ is varied with the temperature fixed until the free energy 
is minimized. The minimum free energy and its corresponding number density 
profile are then the equilibrium values for that temperature.

\section{Soft Fundamental Measure Theory (SFMT)}
In order to use cDFT, a functional for the excess Helmholtz free
energy for the particular fluid being modeled must be formed. But forming 
an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
For the Hard Sphere fluid, there is another way to form an expression for 
the excess Helmholtz free energy in terms of a homogeneous number density 
and, that is to use Scaled Particle Theory (SPT).

Scaled Particle Theory seeks a general expression for the excess chemical 
potential of a homogeneous Hard Sphere fluid which can then be integrated 
to find the excess Helmholtz free energy.
%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogenesous fluid corredsponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemcial potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation is known as Widom's Insertion Formula given by
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
hard sphere, R, and the cavity is either filled or not. \color{black}
%For cavities that are much larger than one of the hard spheres, 
Forming a more general result for different sizes of cavities in the fluid, 
Scaled Partical Theory arrives at an expression for the excess Helmholtz 
free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

@@

@@

\section{Overview}
We use classical Density Functional Theory (cDFT) in conjunction 
with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
energy of a system, and the corresponding equilibrium number density profile. 
In particular, SFMT is used to form the expression for the excess free energy 
without having to take into account the interaction each sphere, or atom, 
has with every other sphere. The excess free energy is then added to the 
ideal free energy calculated analytically to find the resulting free energy 
for the system. 
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
For a spatially varying number density, as would be expected for a crystal, 
the free energy is a functional of the number density profile $n(\vec{r})$  
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}

To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
the temperature fixed until the free energy is minimized. The minimum free 
energy and its corresponding number density profile are then the equilibrium 
values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
the fluid are found, they can be compared to find the resulting state of 
the fluid. Outside the region of coexistence, the fluid will be completely 
in a crystalline state if the crystal free energy is lower than the liquid 
free energy, and it will be completely in a liquid state if the liquid 
free energy is lower than the crystal free energy.

@@

%move this paragraph to methods?
-----------*****-------------

Although the entire crystal lattice can be constructed with FCC cells, 
FCC cells are not the smallest cells with which the lattice can be built. 
The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell} 
is the smallest structure that, when repeated, re-creates the entire crystal 
lattice. The parallepiped primitive cell takes up one-fourth the volume 
of an FCC cell, includes portions of eight atoms for a total of one full 
atom, and has eight lattice sites convenietly reachable by moving along 
one of three different lattice vectors to step from one atom to the next. 
The three lattice vectors will be useful in creating the lattice structure 
computationally, and will also make the program run faster.

----------****--------------
%-------------------------------

  \begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell}
  \end{figure}

@@

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

A simplified flowchart of the program created to investigate the freezing 
of a WCA fluid is shown on %pages x-x. 
The program takes four inputs: reduced number density n*, reduced temperature t*, 
a range of Gaussian 
widths gw, and a range of fraction of vacancies fv.The number density 
profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
is calculated for each profile until the minimum free energy is found for 
that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
The process is then repeated to find the free energies at other number 
densities and temperatures.

@@

*****  end  stuff?  ******

\section{Computing the Liquid and Crystal Ideal Helmholtz free energies}
